Human Pose Estimation for Video Game Control

### 2022-05-07 paper/paper.tex
Added feature classification for steering

With a 128-core NVIDIA GPU on a single-board computer footprint, the Jetson Nano is a popular 
choice for deep model inference on embedded computers. Perhaps more important than the hardware 
itself, however, is the specific Tensor RT optimized models available for use on the Jetson Nano. 
Through a number of optimization techniques such as model quantization, unuzed output elimination 
and horizontal layer fusion/layer aggregation, the Tensorflow with Tensor RT (TF-TRT) backend is 
able to greatly improve inference time on the Jetson Nano. TF-TRT is able to output optimized 
models from Open Neural Network Exchange Format (ONXX) and for popular tasks such as human pose 
estimation, ONXX model formats are readily available from NVIDIA Developers.

\subsection{Pose Estimation}
Using a variant of PoseNet built on a ResNet-18 backbone, we observe 15-17 Frames Per Second 
(FPS) on the Jetson Nano. The so called Pose-ResNet-18-body outputs 17 detected keypoints in 
image coordinate format. 15-17 FPS performance is smooth enough keypoint detection to allow us to 
play these simple games in real time.

\subsection{Feature Tracking and Classification}
\subsubsection{Feature Classification For Steering}
\subsubsection{Feature Classification For Jumping/Ducking}

>>>>

With a 128-core NVIDIA GPU on a single-board footprint, the Jetson Nano is a popular 
choice for deep model inference on embedded computers. Perhaps more important than the hardware 
itself, however, is the specific Tensor RT optimized models available for use on the Jetson Nano. 
Through a number of optimization techniques such as model quantization, unuzed output elimination 
and horizontal layer fusion/layer aggregation, the Tensorflow with Tensor RT (TF-TRT) backend is 
able to greatly improve inference time on the Jetson Nano. TF-TRT is able to output optimized 
models from Open Neural Network Exchange Format (ONXX) and for popular tasks such as human pose 
estimation, ONXX model formats are readily available from NVIDIA Developers.\\

Using a variant of PoseNet built on a ResNet-18 backbone, we observe 15-17 Frames Per Second 
(FPS) on the Jetson Nano. The so called Pose-ResNet-18-body outputs 17 detected keypoints in 
image coordinate format. 15-17 FPS performance is smooth enough keypoint detection to allow us to 
play games in real time.

\subsection{Pose Estimation}
After testing the variety of available solutions for human pose estimation 
above, we identify that running TRT optimized PoseNet through the Jetson Nano 
is the best solution for our application.\\

To build the first phase of our human-computer interaction device, we use the 
jetson-inference and jetson-utils python packages (citation here) 
to build a human pose estimator in Python. The output of this step is a 
collection of 2D keypoints, correpsonding to image coordinates of the locations 
of the 17 body parts detected by PoseNet.

\subsection{Feature Tracking and Classification}
The second major phase of the human-computer interaction device are the 
keypoint classifiers, which receive 2d image coordinates keypoint detections 
from the pose estimator, and classify collections of keypoints into computer 
commands in a game-specific way. In a fully generalized model of our 
human-computer interaction device, the keypoint "decoder" built in this step 
need not even be a classifier. For more complicated continuous inputs to 
video games, this can be a more complicated regression problem as we would 
seek to estimate continuous output based on the locations of human pose 
keypoints.\\

We construct two distinct keypoint classifiers, one for each video game we 
demonstrate our system on. We construct one classifier to derive a steering 
command from pose keypoints, and one to derive jumping/ducking information 
from an identical set of keypoints.

\subsubsection{Feature Classification For Steering}
The first of two example video games we play as demonstrations of our 
human-computer interface is a simple car-racing game, in which the player's 
vehicle must dodge obstacles by steering left and right. The video game 
accepts controls as either left arrow key or right arrow key, so our 
keypoint classifier for this game seeks to classify collections of keypoints 
derived from our human pose estimator into either steer left or steer right 
commands.\\

To maintain a physical intuition to the gestures that are mapped to either 
left arrow key or right arrow key, we require a user to steer the car with 
an imaginary steering wheel on the video stream. From each frame of the user's 
hands on an imaginary wheel, we extract classifications for the game control 
as follows:

\begin{enumerate}
    \item Assert that both the left wrist and right wrist of the user on the 
        imaginary wheel are present in the frame. If one or more wrists are 
        not present in the frame, skip the remaining steps and output no 
        steering command.
    \item Compute the vector connecting the left wrist of the user to the 
        right wrist of the user.
    \item Compute the angle to the horizontal of the vector computed in step 2.
    \item Threshold on the angle to the horizontal with an experimentally 
        determined cutoff threshold.
\end{enumerate}

Let the location (in image coordinates) of the user's left wrist be $(x,y)$ and 
the location of the user's right wrist be $(x', y')$. The displacement vector 
connecting the two wrists is given by:

$$d = (x - x', y - y')$$

The angle to the horizontal is then given by:

$$\theta = \arctan{\frac{y-y'}{x-x'}}$$

After some experimentation with the system, we found that a reasonable threshold 
for the steering angle is $0.2$ radians. Our classifications for steering angle 
are then derived as:

\begin{itemize}
    \item $\theta > 0.2$: Steer left
    \item $-0.2 < \theta < 0.2$: No command
    \item $-0.2 < \theta$: Steer Right
\end{itemize}

\subsubsection{Feature Classification For Jumping/Ducking}
